### 2017-02-05 csf-2017/paper.tex

\section{Introduction}
\label{sec:introduction}

>>>>

\section{Introduction}
\label{sec:introduction}
%%% a)
% Current low-level protection coarse grained memory protection or
% properties of high-level languages are not really enforced?

% Capability machines offer fine-grained memory protection
A capability machine is a low-level machine that offers fine-grained
memory protection.
% Capability machines and unforgeable tokens for memory access. 
This is done by introducing unforgeable tokens to
the machine. These tokens are called capabillities and grant a kind of
authority and a range of authority. We will refer to the capabilities
that grant a combination of read, write, or execute permission as memory
capabilities. Capabilities with the necessary permissions must be
presented every time an instruction that manipulates the memory is
executed.

Memory capabilities provide memory protection, but they do not provide
a way to set up security domains. When you jump to an untrusted piece
of code that is supposed to jump back to you, then you have to either
revoke all your capabilities or pass to the piece of code you do not
trust. This short coming can be overcome by adding something like the
enter capability from the M-Machine\todo{add reference}. The enter
capability is an opaque capability which can only be used for a
jump. When jumped to it grants read and execute permission which
allows the code that is now executing to read capabilities stored in
memory. 
The CHERI processor's ccall achieves a similar property\todo{Add reference}.

Capabilities are irrevocable, so when we pass a capability has to
an untrusted program, then we have to assume that the program keeps it
around indefenitely. This means that we cannot reuse the piece of
memory that the capability governs essentially creating a memory
leak. The CHERI processor has a special kind of capabilities called
\emph{local capabilities}\todo{reference?} that introduces a simple
kind of temporal information control which allows for a simple kind of
capability revocation. This is achieved by adding a tag to every
capability that marks it as either local or global. Local capabilities
can only be written through a capability with a new permission called
\emph{permit write local}.
% More about local capabilities.

With just memory capabilities, enter capabailities, and local
capabilities, the capability machine is very powerful and can
enforce properties that high-level languages promise. In most
high-level languages, we expect local state to be private and
encapsulated. The memory capabilities make sure that local state
cannot be accessed without a capability to do so. With the enter
capabaility, we can regain control of local state after passing
control to an untrusted piece of code which gives us encapsulation
when control is passed to an untrusted program.

Many high-level languages guarantee that function function calls are
well-bracketed. This can easily be achieved by having a trusted call
stack. On a capability machine, it is possible to enforce well
barcketedness without a trusted stack. When ensuring well-bracketed
the two main challenges calls are 1) preventing storage of return
pointers and 2) encapsulation of the local stack frame. If we don't
have the first property, then an adversary can store the return
capability in a call and jump to it in a later call. If the local
stack frame is not properly encapsulated, then an adversary can break
the well-bracketedness by jumping to a different calls return
capability.

In order to deal with the two challenges, we make sure that all return
capabilities are local and that there are no global capabilities with
permit write local permission. This makes sure that there is no way to
save a return capability in persistent local state.\lau{this does not really make
sense as the local stack frame can be seen as local state} It is,
however, to limiting to not provide any means to store the return
capabilities as a piece of code may need to store a return capability
before jumping to an adversary. It is therefore needed to provide a
local capability with permit write local permission which can be used
as a stack. The stack should now be the only place one can store the
return capability, but it should also be the only place one can store
the stack capability.

Using local return capabilities and putting a stack abstraction on a
local capability permit write local capability is a good start, but as
it turns out, it is not wnough. An adversary could do the following 
\begin{enumerate}
\item In a first invocation, fill the stack frame with copies of the
  current stack capability.
\item In a later call, the adversary may be able to load the stack
  capability from the previous call.
\item If we have used a larger part of the stack than in the first
  call, then the adversary has access to part of our local stack frame.
\end{enumerate}

The above attack shows that it is tricky to ensure
well-bracketedness. It raises the question whether it is possible to
do so it is completely watertight. The attack utilized the stack, so
it is necessary to have a trusted stack or can we do without? In
general it raises the question, how do we reason about capability
machines and local capabilities, so we are sure that what ever calling
scheme we come up with actually provably ensures well-bracketed calls.

In the following paper, we present the following contributions
\begin{enumerate}
\item 
\end{enumerate}

%%% b)

% Examples of capability machines?

% Enter capabilities

% Local capabilities

% High-level languages often provide guarantees such as encapsulation of private state and well-bracketedness of calls - often not clear how well this is enforced. Just enforced when interacting with other programs written in the same high-level language or is it also guaranteed when interacting with assembly programs.

% High-level programs compiled to assembly can ensure these properties.

%

### 2017-02-05 local-capability-machine.tex

\begin{proof}[Proof of Theorem~\ref{thm:world-existence}]
  The theorem follows from a more general solution theorem for the
  category of $P$ preordered \cofes, see~\cite{bizjak-note}. 
  We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to
  $P$. 
  \begin{align*}
     F_1((X,{\futurestr}'),(Y,{\futurewk}')) = \\
     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,{\futurewk}') \monnefun \UPred{\HeapSegments})) + \\
                                          & \{\perma\} \times \States
                                            \times \Rels \times
                                            (\States \fun ((X,{\futurestr}')
                                            \monnefun
                                            \UPred{\HeapSegments})))),
                                        \futurestr)
  \end{align*}
  and
  \begin{align*}
     F_2((X,{\futurestr}'),(Y,{\futurewk}')) = \\
     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,{\futurewk}') \monnefun \UPred{\HeapSegments})) + \\
                                          & \{\perma\} \times \States
                                            \times \Rels \times
                                            (\States \fun ((X,{\futurestr}')
                                            \monnefun
                                            \UPred{\HeapSegments})))),
                                        \futurestr)
  \end{align*}
  The orderings $\futurestr$ and $\futurewk$ used in the definition of
  $F_1$ and $F_2$ are defined by the properties given above.
  Note that the image of $F_1$ and $F_2$ only differ in the ordering
  relation, i.e., letting $U$ denote the forgetful functor from the
  category of preordered \cofes\ to the category of \cofes, we have
  $U \circ F_1 = U \circ F_2$.
  From~\cite{bizjak-note} it then follows that there exists a \cofe\ $\Wor$ and two
  preorderings $\futurestr$ and $\futurewk$ and an isomorphism $\xi$ satisfying the properties
  claimed in theorem.
  (Here, in the proof, we have written the ordering explicitly on the \cofe\ when
  using monotone non-expansive functions; in the theorem formulation
  we have instead annotated the arrow to indicate which ordering is used.)
\end{proof}

>>>>

\begin{proof}[Proof of Theorem~\ref{thm:world-existence}]
  The theorem follows from a more general solution theorem for the
  category of $P$ preordered \cofes,
  see~\cite{Birkedal:2010:TCS:411:4102-4122}, \cite{Birkedal:tutorial-notes} and \cite{bizjak-note}. 
  We define two functors $F_1$ and $F_2$ from $P^{op}\times P^{op}$ to
  $P$. 
  \begin{align*}
     F_1((X,{\futurestr}'),(Y,{\futurewk}')) = \\
     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,{\futurewk}') \monnefun \UPred{\HeapSegments})) + \\
                                          & \{\perma\} \times \States
                                            \times \Rels \times
                                            (\States \fun ((X,{\futurestr}')
                                            \monnefun
                                            \UPred{\HeapSegments})))),
                                        \futurestr)
  \end{align*}
  and
  \begin{align*}
     F_2((X,{\futurestr}'),(Y,{\futurewk}')) = \\
     (\blater (\nats \finparfun ( & \{\revoked\}  + \\
                                          & \{\temp\} \times \States \times \Rels \times (\States \fun ((Y,{\futurewk}') \monnefun \UPred{\HeapSegments})) + \\
                                          & \{\perma\} \times \States
                                            \times \Rels \times
                                            (\States \fun ((X,{\futurestr}')
                                            \monnefun
                                            \UPred{\HeapSegments})))),
                                        \futurewk)
  \end{align*}
  The orderings $\futurestr$ and $\futurewk$ used in the definition of
  $F_1$ and $F_2$ are defined by the properties given above.
  Note that the image of $F_1$ and $F_2$ only differ in the ordering
  relation, i.e., letting $U$ denote the forgetful functor from the
  category of preordered \cofes\ to the category of \cofes, we have
  $U \circ F_1 = U \circ F_2$.
  From~\cite{bizjak-note} it then follows that there exists a \cofe\ $\Wor$ and two
  preorderings $\futurestr$ and $\futurewk$ and an isomorphism $\xi$ satisfying the properties
  claimed in theorem.
  (Here, in the proof, we have written the ordering explicitly on the \cofe\ when
  using monotone non-expansive functions; in the theorem formulation
  we have instead annotated the arrow to indicate which ordering is used.)
\end{proof}
